---
title: "Single_Cell_CAOV3_multiplexed"
output: html_document
date: "2023-05-04"
---
For this experiment, we are looking at the data from the multiplexed time points 
from CAOV3. The time points included are P0, P1, P6, P7.  They are sequenced where each should have 5k cells. Actual Totals according to cell ranger were: 
#P0
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
```

```{r}
matrix_dir = "C:/Users/empage/Documents/per_sample_outs/p0/count/sample_feature_bc_matrix/"
barcode.path <- paste0(matrix_dir, "barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "features.tsv.gz")
matrix.path <- paste0(matrix_dir, "matrix.mtx.gz")
mat <- Matrix::readMM(file = matrix.path)
feature.names = read.delim(features.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(mat) = barcode.names$V1
rownames(mat) = feature.names$V1
```

```{r}
pbmc <- CreateSeuratObject(counts = mat, min.cells = 1, min.features = 200)
pbmc
```

```{r}
#add meta separately 
x1 <- rep("P0", each= 2253)
pbmc <- AddMetaData(object= pbmc, metadata = x1, col.name= "time")
```

```{r}
# Visualize QC metrics as a violin plot
VlnPlot(pbmc, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by= )
```
```{r}
# FeatureScatter is typically used to visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.

plot2 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2
```
Trying without subsetting because we have a limied number of cells
```{r}
#pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500)
```


```{r}
pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
pbmc <- FindVariableFeatures(pbmc, selection.method = "mvp", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(pbmc), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(pbmc)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1
```
```{r}
plot2
```

```{r}
all.genes <- rownames(pbmc)
pbmc <- ScaleData(pbmc, features = all.genes)
```

```{r}
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
```

```{r}
print(pbmc[["pca"]], dims = 1:5, nfeatures = 5)
```

```{r}
VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")
```

```{r}
DimPlot(pbmc, reduction = "pca", group.by= )
```

```{r}
DimHeatmap(pbmc, dims = 1:5, cells = 500, balanced = TRUE)
```
```{r}
DimHeatmap(pbmc, dims = 6:10, cells = 500, balanced = TRUE)
```

```{r}
DimHeatmap(pbmc, dims = 11:15, cells = 500, balanced = TRUE)
```
```{r}
DimHeatmap(pbmc, dims = 16:20, cells = 500, balanced = TRUE)
```
```{r}
DimHeatmap(pbmc, dims = 21:25, cells = 500, balanced = TRUE)
```


#editing this to reflect that higher dimensions and a lower resolution need to be used to reflect the data better
```{r}
pbmc <- FindNeighbors(pbmc, dims = 1:20)
pbmc <- FindClusters(pbmc, resolution = 0.5)
```
#13 dims seems to be where it starts to keep the same for higher dimensions, so this is what I chose 
```{r}
pbmc <- RunUMAP(pbmc, dims = 1:20)
DimPlot(pbmc, reduction = "umap")
```
```{r}
pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
pbmc.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)

write.csv(pbmc.markers, "C:/Users/empage/Documents/P0_genes_subgroups_no_subset.csv")
```
#P1 
```{r}
matrix_dir = "C:/Users/empage/Documents/per_sample_outs/p1/count/sample_feature_bc_matrix/"
barcode.path <- paste0(matrix_dir, "barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "features.tsv.gz")
matrix.path <- paste0(matrix_dir, "matrix.mtx.gz")
mat <- Matrix::readMM(file = matrix.path)
feature.names = read.delim(features.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(mat) = barcode.names$V1
rownames(mat) = feature.names$V1
```

```{r}
pbmc_1 <- CreateSeuratObject(counts = mat, min.cells = 1, min.features = 200)
pbmc_1
```
```{r}
#add meta separately 
x2 <- rep("P1", each= 897)
pbmc_1 <- AddMetaData(object= pbmc_1, metadata = x2, col.name= "time")
```

```{r}
# Visualize QC metrics as a violin plot
VlnPlot(pbmc_1, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by= )
```

```{r}
# FeatureScatter is typically used to visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.
plot2 <- FeatureScatter(pbmc_1, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2
```

```{r}
#pbmc_1 <- subset(pbmc_1, subset = nFeature_RNA > 200 & nFeature_RNA < 2500)
```


```{r}
pbmc_1 <- NormalizeData(pbmc_1, normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
pbmc_1 <- FindVariableFeatures(pbmc_1, selection.method = "mvp", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(pbmc_1), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(pbmc_1)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1
```
```{r}
plot2
```

```{r}
all.genes <- rownames(pbmc_1)
pbmc_1 <- ScaleData(pbmc_1, features = all.genes)
```

```{r}
pbmc_1 <- RunPCA(pbmc_1, features = VariableFeatures(object = pbmc_1))
```

```{r}
print(pbmc_1[["pca"]], dims = 1:5, nfeatures = 5)
```

```{r}
VizDimLoadings(pbmc_1, dims = 1:2, reduction = "pca")
```

```{r}
DimPlot(pbmc_1, reduction = "pca", group.by= )
```

```{r}
DimHeatmap(pbmc_1, dims = 1, cells = 500, balanced = TRUE)
```

```{r}
pbmc_1 <- FindNeighbors(pbmc_1, dims = 1:20)
pbmc_1 <- FindClusters(pbmc_1, resolution = 0.5)
```

```{r}
pbmc_1 <- RunUMAP(pbmc_1, dims = 1:20)
DimPlot(pbmc_1, reduction = "umap")
```

```{r}
pbmc.markers <- FindAllMarkers(pbmc_1, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
pbmc.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
write.csv(pbmc.markers, "C:/Users/empage/Documents/P1_genes_subgroups_no_subset.csv")
```
#P6
```{r}
matrix_dir = "C:/Users/empage/Documents/per_sample_outs/p6/count/sample_feature_bc_matrix/"
barcode.path <- paste0(matrix_dir, "barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "features.tsv.gz")
matrix.path <- paste0(matrix_dir, "matrix.mtx.gz")
mat <- Matrix::readMM(file = matrix.path)
feature.names = read.delim(features.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(mat) = barcode.names$V1
rownames(mat) = feature.names$V1
```

```{r}
pbmc_3 <- CreateSeuratObject(counts = mat, min.cells = 1, min.features = 200)
pbmc_3
```
```{r}
#add meta separately 
x3 <- rep("P6", each= 1794)
pbmc_3 <- AddMetaData(object= pbmc_3, metadata = x3, col.name= "time")
```

```{r}
# Visualize QC metrics as a violin plot
VlnPlot(pbmc_3, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by= )
```

```{r}
# FeatureScatter is typically used to visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.
plot2 <- FeatureScatter(pbmc_3, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2
```

```{r}
#pbmc_3 <- subset(pbmc_3, subset = nFeature_RNA > 200 & nFeature_RNA < 2500)
```

```{r}
pbmc_3 <- NormalizeData(pbmc_3, normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
pbmc_3 <- FindVariableFeatures(pbmc_3, selection.method = "mvp", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(pbmc_3), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(pbmc_3)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1
```

```{r}
plot2
```

```{r}
all.genes <- rownames(pbmc_3)
pbmc_3 <- ScaleData(pbmc_3, features = all.genes)
```

```{r}
pbmc_3 <- RunPCA(pbmc_3, features = VariableFeatures(object = pbmc_3))
```

```{r}
print(pbmc_3[["pca"]], dims = 1:5, nfeatures = 5)
```

```{r}
VizDimLoadings(pbmc_3, dims = 1:2, reduction = "pca")
```

```{r}
DimPlot(pbmc_3, reduction = "pca", group.by= )
```

```{r}
DimHeatmap(pbmc_3, dims = 1, cells = 500, balanced = TRUE)
```

#also edited to higher dims (from 3 to 13) and chnage in resolution (from 0.5 to 0.25)
```{r}
pbmc_3 <- FindNeighbors(pbmc_3, dims = 1:20)
pbmc_3 <- FindClusters(pbmc_3, resolution = 0.5)
```

```{r}
pbmc_3 <- RunUMAP(pbmc_3, dims = 1:20)
DimPlot(pbmc_3, reduction = "umap")
```

```{r}
pbmc.markers <- FindAllMarkers(pbmc_3, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
pbmc.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
write.csv(pbmc.markers, "C:/Users/empage/Documents/P6_genes_subgroups_no_subset.csv")
```

#P7
```{r}
matrix_dir <- "C:/Users/empage/Documents/per_sample_outs/p7/count/sample_feature_bc_matrix/"
barcode.path <- paste0(matrix_dir, "barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "features.tsv.gz")
matrix.path <- paste0(matrix_dir, "matrix.mtx.gz")
mat <- Matrix::readMM(file = matrix.path)
feature.names = read.delim(features.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(mat) = barcode.names$V1
rownames(mat) = feature.names$V1
```

```{r}
pbmc_4 <- CreateSeuratObject(counts = mat, min.cells = 1, min.features = 200)
pbmc_4
```

```{r}
#add meta separately 
x4 <- rep("P7", each= 1617)
pbmc_4 <- AddMetaData(object= pbmc_4, metadata = x4, col.name= "time")
```

```{r}
# Visualize QC metrics as a violin plot
VlnPlot(pbmc_4, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by= )
```

```{r}
# FeatureScatter is typically used to visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.
plot2 <- FeatureScatter(pbmc_4, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2
```

```{r}
#pbmc_4 <- subset(pbmc_4, subset = nFeature_RNA > 200 & nFeature_RNA < 2500)
```

```{r}
pbmc_4 <- NormalizeData(pbmc_4, normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
pbmc_4 <- FindVariableFeatures(pbmc_4, selection.method = "mvp", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(pbmc_4), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(pbmc_4)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1
```

```{r}
plot2
```

```{r}
all.genes <- rownames(pbmc_4)
pbmc_4 <- ScaleData(pbmc_4, features = all.genes)
```

```{r}
pbmc_4 <- RunPCA(pbmc_4, features = VariableFeatures(object = pbmc_4))
```

```{r}
print(pbmc_4[["pca"]], dims = 1:5, nfeatures = 5)
```

```{r}
VizDimLoadings(pbmc_4, dims = 1:2, reduction = "pca")
```

```{r}
DimPlot(pbmc_4, reduction = "pca", group.by= )
```

```{r}
DimHeatmap(pbmc_4, dims = 1, cells = 500, balanced = TRUE)
```

#also edited to higher dims (from 3 to 13) and chnage in resolution (from 0.5 to 0.25)
```{r}
pbmc_4 <- FindNeighbors(pbmc_4, dims = 1:20)
pbmc_4 <- FindClusters(pbmc_4, resolution = 0.5)
```

```{r}
pbmc_4 <- RunUMAP(pbmc_4, dims = 1:20)
DimPlot(pbmc_4, reduction = "umap")
```

```{r}
pbmc.markers <- FindAllMarkers(pbmc_4, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
pbmc.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
write.csv(pbmc.markers, "C:/Users/empage/Documents/P7_genes_subgroups_no_subset.csv")
```

#Merging 
```{r}
pbmc.combined <- merge(pbmc, y = c(pbmc_1, pbmc_3, pbmc_4), add.cell.ids = c("P0","P1","P6","P7"), project = "CAOV3_multiplexed")
pbmc.combined
```
```{r}
pbmc.combined <- merge(pbmc, y = c(pbmc_1, pbmc_3, pbmc_4), add.cell.ids = c("P0","P1","P6","P7"), project = "CAOV3_multiplexed")
pbmc.combined
```

```{r}
# Visualize QC metrics as a violin plot
VlnPlot(pbmc.combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by= )
```

```{r}
# FeatureScatter is typically used to visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.
plot2 <- FeatureScatter(pbmc.combined, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2
```

```{r}
#pbmc.combined <- subset(pbmc.combined , subset = nFeature_RNA > 200 & nFeature_RNA < 2500)

pbmc.combined  <- NormalizeData(pbmc.combined , normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
pbmc.combined <- FindVariableFeatures(pbmc.combined, selection.method = "mvp", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(pbmc.combined), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(pbmc.combined)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1
```

```{r}
plot2
```

```{r}
all.genes <- rownames(pbmc.combined)
pbmc.combined <- ScaleData(pbmc.combined, features = all.genes)
```

```{r}
pbmc.combined <- RunPCA(pbmc.combined, features = VariableFeatures(object = pbmc.combined))
```

```{r}
print(pbmc.combined[["pca"]], dims = 1:5, nfeatures = 5)
```

```{r}
VizDimLoadings(pbmc.combined, dims = 1:2, reduction = "pca")
```

```{r}
DimPlot(pbmc.combined, reduction = "pca", group.by= )
```

```{r}
DimHeatmap(pbmc.combined, dims = 1, cells = 500, balanced = TRUE)
```

```{r}
pbmc.combined <- FindNeighbors(pbmc.combined, dims = 1:20)
pbmc.combined <- FindClusters(pbmc.combined, resolution = 0.5)
```

#Starting the combined exploration here, so there will be a few UMAPs with different dims 
```{r}
pbmc.combined <- RunUMAP(pbmc.combined, dims = 1:20)
DimPlot(pbmc.combined, reduction = "umap")
```

```{r}
DimPlot(object = pbmc.combined, reduction = "umap", group.by = "time")
```

```{r}
pbmc.markers <- FindAllMarkers(pbmc.combined, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
pbmc.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
write.csv(pbmc.markers, "C:/Users/empage/Documents/CAOV3_genes_subgroups_no_subset.csv")
```
