---
title: "UMAP Exploration P7"
output: html_document
date: "2022-11-14"
---
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
```

```{r}
# Load the PBMC dataset
pbmc.data <- Read10X(data.dir = "~/Documents/Mythreye_Lab/Mehri_Resha/Single_cell_mehri/run_count_P7")
# Initialize the Seurat object with the raw (non-normalized data).
pbmc <- CreateSeuratObject(counts = pbmc.data, min.cells = 1, min.features = 200)
pbmc
```

```{r}
#add meta separately 
x2 <- rep("P7", each= 14029)
pbmc <- AddMetaData(object= pbmc, metadata = x2, col.name= "time")
```

```{r}
pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500)

pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
pbmc <- FindVariableFeatures(pbmc, selection.method = "mvp", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(pbmc), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(pbmc)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1
```

```{r}
all.genes <- rownames(pbmc)
pbmc <- ScaleData(pbmc, features = all.genes)
```

```{r}
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
```

```{r}
print(pbmc[["pca"]], dims = 1:20, nfeatures = 5)
```

```{r}
VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")
```

```{r}
DimHeatmap(pbmc, dims = 1, cells = 500, balanced = TRUE)
```

#Round 1 of exploration: Change the dims within the RunUMAP function. 

```{r}
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)
```

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:3)
DimPlot(pbmc, reduction = "umap")
```

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:5)
DimPlot(pbmc, reduction = "umap")
```

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:7)
DimPlot(pbmc, reduction = "umap")
```

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:10)
DimPlot(pbmc, reduction = "umap")
```

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:13)
DimPlot(pbmc, reduction = "umap")
```

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:15)
DimPlot(pbmc, reduction = "umap")
```

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:17)
DimPlot(pbmc, reduction = "umap")
```

From here, I noticed that some of these subclusters just do not want to separate further. So I wanted a UMAP to incorporate that idea. The best way to do that is to reduce the resolution, which is what I did. My expectation is that it will reduce the subclusters and hold on to the most important features. 

#Round 2: Change the Resolution in FindClusters ot 0.25 (half of the original) and explore the different dims in RunUMAP 
```{r}
pbmc2 <- FindNeighbors(pbmc, dims = 1:10)
pbmc2 <- FindClusters(pbmc2, resolution = 0.25)
```

```{r}
pbmc2 <- RunUMAP(pbmc2, dims = 1:3)
DimPlot(pbmc2, reduction = "umap")
```

```{r}
pbmc2 <- RunUMAP(pbmc2, dims = 1:5)
DimPlot(pbmc2, reduction = "umap")
```

```{r}
pbmc2 <- RunUMAP(pbmc2, dims = 1:7)
DimPlot(pbmc2, reduction = "umap")
```

```{r}
pbmc2 <- RunUMAP(pbmc2, dims = 1:10)
DimPlot(pbmc2, reduction = "umap")
```

```{r}
pbmc2 <- RunUMAP(pbmc2, dims = 1:13)
DimPlot(pbmc2, reduction = "umap")
```

```{r}
pbmc2 <- RunUMAP(pbmc2, dims = 1:15)
DimPlot(pbmc2, reduction = "umap")
```

```{r}
pbmc2 <- RunUMAP(pbmc2, dims = 1:17)
DimPlot(pbmc2, reduction = "umap")
```


```{r}
pbmc.markers <- FindAllMarkers(pbmc2, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
pbmc.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)

write.csv(pbmc.markers, "~/Desktop/P7_genes_subgroups_exploration.csv")
```


```{r}
pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
pbmc.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)

write.csv(pbmc.markers, "~/Desktop/P7_genes_subgroups_exploration_original.csv")
```